\section*{Adaptive QRC Roadmap}

\begin{itemize}
    \item What we add: a state-aware instructor model for group synchronization.
    \item Where it enters: as a closed-loop coordinator that reads dispersion, lag, suppression, and contamination.
    \item Why it matters: the coordinator adapts rhythm and focus instead of forcing uniform behavior.
    \item What we keep separate: memory stays in $K$; suppression stays in $S$; coordination stays in RC.
\end{itemize}

\paragraph{1. Core role.}
The adaptive coordinator is not a fixed metronome. It is a soft controller that changes its own gain and focus according to the group response. If dispersion and lag rise, the coordinator strengthens guidance. If the group stabilizes, guidance is relaxed.

\paragraph{2. Variables.}
\begin{itemize}
    \item $S$ --- suppression, isolation, or temporary unreadiness;
    \item $K$ --- experience, bounded memory, consolidation;
    \item $\mathcal{D}$ --- phase dispersion;
    \item $L$ --- lag or delayed adaptation;
    \item $A$ --- stable anchors;
    \item $C$ --- contagion or propagation of instability.
\end{itemize}

\paragraph{3. What each variable does.}
\begin{itemize}
    \item $S$ tells the coordinator which agents are not yet ready for strong coupling.
    \item $K$ records accumulated experience and saturates instead of growing without bound.
    \item $\mathcal{D}$ measures how far the group has drifted from common rhythm.
    \item $L$ indicates which subgroup is slow to adapt.
    \item $A$ identifies stable agents that can serve as anchors.
    \item $C$ flags the risk that a local failure spreads to neighbors.
\end{itemize}

\paragraph{4. Minimal closed-loop update.}
\begin{align}
\phi_{\mathrm{gain}}(t) &= \mathrm{clamp}\!\left(\phi_0\bigl[1+a_1\mathcal{D}(t)+a_2\overline{L}(t)+a_3 C(t)\bigr]\right), \\
\Phi_{t+1} &= \Phi_t + \Delta t\left[-\kappa(\mathcal{D},L)\Phi_t + \mathrm{Arg}\bigl(\langle e^{i\theta}\rangle\bigr)\right].
\end{align}
This is the minimal form of the instructor model: state is observed, gain is adapted, and the global rhythm is shifted rather than enforced.

\paragraph{5. Focus and gating.}
Stable agents should receive more weight, while unstable neighbors should contribute less.
\begin{align}
w_i &\propto (1-S_i), \\
\varepsilon_{ij} &= \varepsilon(1-S_i)(1-S_j).
\end{align}
This implements two rules:
\begin{itemize}
    \item look first at agents that remain stable;
    \item reduce contagion from agents that are still collapsing.
\end{itemize}

\paragraph{6. What to test.}
\begin{itemize}
    \item baseline RC vs adaptive RC;
    \item recovery time under different lag profiles;
    \item whether the adaptive version reduces critical-noise sensitivity;
    \item whether stable anchors improve recovery in mixed-experience groups.
\end{itemize}

\paragraph{7. Writing rule.}
Keep the roles separated:
\begin{itemize}
    \item $S$ is not memory;
    \item $K$ is not control;
    \item RC is not hard forcing;
    \item the adaptive layer only reads state and changes its gain.
\end{itemize}

\selectlanguage{russian}
\section*{Памятка по Adaptive QRC}

\begin{itemize}
    \item Что добавляем: инструктора-координатора, который подстраивает ритм по состоянию группы.
    \item Куда вводим: в замкнутый контур, где координатор читает дисперсию, лаг, подавление и заражение.
    \item Зачем это нужно: чтобы ритм менялся по реакции группы, а не навязывался всем одинаково.
    \item Что держим отдельно: память живёт в $K$, подавление --- в $S$, координация --- в RC.
\end{itemize}

\paragraph{1. Главная роль.}
Адаптивный координатор --- это не фиксированный метроном, а мягкий управляющий слой. Он меняет усиление и фокус по ответу группы. Если дисперсия и лаг растут, он усиливает ведение. Если группа стабилизируется, воздействие ослабляется.

\paragraph{2. Переменные.}
\begin{itemize}
    \item $S$ --- подавление, изоляция или временная неготовность;
    \item $K$ --- опыт, ограниченная память, консолидация;
    \item $\mathcal{D}$ --- фазовая дисперсия;
    \item $L$ --- лаг или запаздывание адаптации;
    \item $A$ --- опорные агенты;
    \item $C$ --- заражение ошибкой или нестабильностью.
\end{itemize}

\paragraph{3. Что делает каждая переменная.}
\begin{itemize}
    \item $S$ показывает, кто ещё не готов к сильной связи.
    \item $K$ накапливает опыт и насыщается, а не растёт бесконечно.
    \item $\mathcal{D}$ показывает, насколько группа разошлась по ритму.
    \item $L$ показывает, какая подгруппа запаздывает.
    \item $A$ указывает на устойчивые элементы, на которые можно опереться.
    \item $C$ показывает риск распространения сбоя.
\end{itemize}

\paragraph{4. Минимальное замкнутое обновление.}
\begin{align}
\phi_{\mathrm{gain}}(t) &= \mathrm{clamp}\!\left(\phi_0\bigl[1+a_1\mathcal{D}(t)+a_2\overline{L}(t)+a_3 C(t)\bigr]\right), \\
\Phi_{t+1} &= \Phi_t + \Delta t\left[-\kappa(\mathcal{D},L)\Phi_t + \mathrm{Arg}\bigl(\langle e^{i\theta}\rangle\bigr)\right].
\end{align}
Это и есть минимальная форма модели инструктора: состояние наблюдается, усиление адаптируется, ритм сдвигается, но не навязывается.

\paragraph{5. Фокус и gate.}
Стабильным агентам даём больший вес, а нестабильные соседи должны влиять слабее.
\begin{align}
w_i &\propto (1-S_i), \\
\varepsilon_{ij} &= \varepsilon(1-S_i)(1-S_j).
\end{align}
Это реализует два правила:
\begin{itemize}
    \item сначала смотрим на устойчивых;
    \item не даём сбою заражать соседей без фильтра.
\end{itemize}

\paragraph{6. Что проверять.}
\begin{itemize}
    \item базовый RC против adaptive RC;
    \item время восстановления при разных профилях лага;
    \item снижает ли адаптация чувствительность к критическому шуму;
    \item помогают ли опорные агенты в смешанных группах.
\end{itemize}

\paragraph{7. Правило письма.}
Роли должны быть разделены:
\begin{itemize}
    \item $S$ не является памятью;
    \item $K$ не является управлением;
    \item RC не является жёстким принуждением;
    \item адаптивный слой только читает состояние и меняет своё усиление.
\end{itemize}
\selectlanguage{english}
