\section{Group Crisis Score}
For the group-level memory, the crisis score should reflect whether an episode was structurally informative rather than merely noisy. A practical score can be built from four normalized components:
\begin{equation}
C_E = w_s\,\tilde S_{\max,E} + w_r\,\tilde R_E + w_t\,\tilde T_E + w_d\,\tilde D_E,
\end{equation}
where:
\begin{itemize}
    \item $\tilde S_{\max,E}$ is the normalized peak mean suppression during episode $E$,
    \item $\tilde R_E$ is the normalized recovery time,
    \item $\tilde T_E$ is the normalized tail span of agent recovery times,
    \item $\tilde D_E$ is the normalized phase dispersion during the episode.
\end{itemize}
All terms are scaled to $[0,1]$ over the current sweep or dataset so that the score remains bounded and comparable across episodes.

The gate is then defined by
\begin{equation}
q_E = \Theta(C_E - C_{\mathrm{crit}}),
\end{equation}
with a practical starting range
\begin{equation}
C_{\mathrm{crit}} \in [0.55, 0.65].
\end{equation}
In the first implementation, $C_{\mathrm{crit}} \approx 0.60$ is a reasonable default: it is high enough to ignore noise, but low enough to retain structurally meaningful crisis episodes.

\selectlanguage{russian}
\section{Групповой кризисный скор}
Для групповой памяти кризисный скор должен отражать не любой шум, а только эпизод, который действительно чему-то учит. Практический скор можно собрать из четырех нормированных компонент:
\begin{equation}
C_E = w_s\,\tilde S_{\max,E} + w_r\,\tilde R_E + w_t\,\tilde T_E + w_d\,\tilde D_E,
\end{equation}
где:
\begin{itemize}
    \item $\tilde S_{\max,E}$ — нормированный пик среднего подавления в эпизоде $E$,
    \item $\tilde R_E$ — нормированное время восстановления,
    \item $\tilde T_E$ — нормированный хвост времен восстановления агентов,
    \item $\tilde D_E$ — нормированная фазовая дисперсия в эпизоде.
\end{itemize}
Все компоненты масштабируются в $[0,1]$ на текущем наборе прогона или свипа, чтобы скор оставался ограниченным и сопоставимым между эпизодами.

Дальше gate задается как
\begin{equation}
q_E = \Theta(C_E - C_{\mathrm{crit}}),
\end{equation}
с практическим стартовым диапазоном
\begin{equation}
C_{\mathrm{crit}} \in [0.55, 0.65].
\end{equation}
Для первой версии разумный дефолт — $C_{\mathrm{crit}} \approx 0.60$: это достаточно высоко, чтобы отсечь шум, и достаточно низко, чтобы не потерять значимые кризисные эпизоды.
\selectlanguage{english}
