\documentclass[12pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[russian,english]{babel}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{graphicx}
\geometry{margin=2.5cm}

\title{
    \textbf{Meta-Stable Architectures}\\
    \Large Volume V (draft): Adaptive QRC (Instructor Model)
}
\author{Symbiosis: Katia \& GPT-5 (Jippi)}
\date{2026}

\begin{document}
\maketitle

\begin{abstract}
We introduce an Adaptive QRC (Instructor Model) that modulates network synchronization based on real-time group state. Unlike a fixed metronome, the adaptive coordinator estimates lagging subgroups, local instability, and coherence dispersion, then updates coupling intensity and focus gates. The objective is to preserve a shared rhythm without enforcing uniform behavior.
\end{abstract}

\selectlanguage{russian}
\begin{abstract}
Мы вводим адаптивный QRC (Instructor Model), который модулирует синхронизацию сети на основе текущего состояния группы. В отличие от фиксированного метронома, координатор оценивает запаздывающие подгруппы, локальную нестабильность и дисперсию когерентности, после чего обновляет силу связи и фокусные гейты. Цель — сохранять общий ритм без жёсткого выравнивания поведения.
\end{abstract}
\selectlanguage{english}

\section{Concept}
The Instructor Model treats the coordinator as a soft, state-aware conductor. It adapts the rhythm and coupling based on observed lag, crisis contagion, and coherence dispersion. The mechanism is informational rather than forceful: the system re-aligns by recognizing patterns of stable agents and protecting unstable agents from contagion.

\selectlanguage{russian}
\section{Концепция}
Instructor Model трактует координатор как мягкого, чувствительного к состоянию дирижёра. Он адаптирует ритм и связи, опираясь на запаздывание, кризисное заражение и дисперсию когерентности. Механизм информационный, а не силовой: система перенастраивается, ориентируясь на устойчивые паттерны и защищая нестабильные элементы от заражения.
\selectlanguage{english}

\section{Adaptive QRC Dynamics}
Let $i=1,\dots,N$ denote agents with phases $\theta_i$ and suppression $S_i$. Define dispersion
\begin{equation}
\mathcal{D}(t)=1-\left|\langle e^{i\theta}\rangle\right|,
\end{equation}
and a lag indicator $L_i$ (local recovery delay or slow subgroup membership). The adaptive coordinator updates its gain:
\begin{equation}
\phi_{\mathrm{gain}}(t)=\phi_0\left[1+\alpha_1\,\mathcal{D}(t)+\alpha_2\,\overline{L}(t)\right],
\end{equation}
with bounds $\phi_{\min}\le\phi_{\mathrm{gain}}\le\phi_{\max}$.

\paragraph{Weighted coupling (focus to stable agents).}
Let weights $w_i\propto (1-S_i)$ emphasize stable agents. Then the coupling term becomes
\begin{equation}
\dot{y}_i\leftarrow \dot{y}_i + \varepsilon(1-S_i)\sum_{j}\,w_j\,(y_j-y_i).
\end{equation}
This implements "look at the experienced neighbor" behavior.

\paragraph{Anti-contagion gate.}
Suppress influence from unstable neighbors via
\begin{equation}
\varepsilon_{ij}=\varepsilon\,(1-S_i)(1-S_j),
\end{equation}
so contagion from high $S_j$ is reduced.

\section{Adaptive Rhythm}
The metronome itself adapts to group response:
\begin{equation}
\dot{\Phi}=-\kappa\Phi + \mathrm{Arg}\big(\langle e^{i\theta}\rangle\big),
\end{equation}
with $\kappa$ modulated by lag and dispersion. This shifts the rhythm rather than forcing it.

\section{Purpose}
Adaptive QRC aims to:
\begin{itemize}
    \item keep global rhythm while preserving local diversity,
    \item reduce contagion by directing attention to stable subgroups,
    \item lower recovery time without rigid control.
\end{itemize}

\selectlanguage{russian}
\section{Цель}
Adaptive QRC предназначен для:
\begin{itemize}
    \item сохранения общего ритма при локальном разнообразии,
    \item снижения заражения через фокус на устойчивых подгруппах,
    \item уменьшения времени восстановления без жёсткого управления.
\end{itemize}
\selectlanguage{english}


\section{Minimal Pseudocode}
\begin{verbatim}
for each step:
  observe phase dispersion D and lag L
  update phi_gain = clamp(phi0*(1+a1*D + a2*L))
  update Phi by mean phase (soft metronome)
  weight neighbors by (1 - S)
  apply coupling and anti-contagion gate
\end{verbatim}

\section{Relation to Vol IV}
This model extends the fixed QRC layer from Volume IV by making its parameters adaptive to group state. It preserves the same soft-control philosophy while adding instructor-like feedback.

\selectlanguage{russian}
\section{????? ? ????? IV}
?????? ????????? QRC ?? ???? IV: ????????? ?????????? ??????????? ? ??????? ?? ????????? ??????. ??? ???? ??????????? ?????? ????????? ????????.
\selectlanguage{english}

\input{design_memo_adaptive_qrc.tex}


\section{Quantum QRC (Conceptual Bridge)}
We propose a quantum analogue of recognition and coherence-driven relaxation using a shared resonator as a soft metronome and sensor.

\paragraph{Resonator phase as metronome.}
Let $a$ be the cavity field. Define the global phase
\begin{equation}
\Phi = \arg\langle a\rangle.
\end{equation}
This provides a weak, global reference without projective measurement of individual phases.

\paragraph{Recognition without collapse.}
In dispersive coupling, a weak probe acquires a phase shift that depends on the qubit state. The effective match can be written as
\begin{equation}
\mathrm{match}_i \propto \mathrm{Re}\left(\langle \sigma_+^i\, a\rangle e^{-i\Phi}\right),
\end{equation}
which acts as a soft coherence indicator rather than a hard measurement.

\paragraph{Coherence-driven relaxation.}
Define dispersion via the resonator field,
\begin{equation}
\mathcal{D}=1-\left|\langle a\rangle\right|.
\end{equation}
Then a soft relaxation channel can be written as
\begin{equation}
\dot S_i \leftarrow \dot S_i - \gamma_{coh}(1-\mathcal{D})S_i.
\end{equation}
This yields recovery without strong forcing, analogous to the classical QRC mechanism.

\paragraph{Why this is new.}
The novelty is the combination: (i) soft metronome (resonator phase), (ii) recognition via weak phase shift, and (iii) coherence-driven relaxation, which together enable recovery from trapped states without hard resets.

\selectlanguage{russian}
\section{????????? QRC (?????????????? ????)}
?? ?????????? ????????? ?????? ????????????? ? ??????????? ?????????? ????? ????? ????????? ??? ?????? ???????? ? ??????.

\paragraph{???? ?????????? ??? ????????.}
????? $a$ ? ???? ??????????. ????????? ?????????? ????
\begin{equation}
\Phi = \arg\langle a\rangle.
\end{equation}
??? ???? ?????? ?????????? ?????? ??? ???????????? ????????? ??? ????????? ???????.

\paragraph{????????????? ??? ????????.}
? ???????????? ????? ?????? ???? ???????? ??????? ?????, ????????? ?? ????????? ??????. ??????????? match:
\begin{equation}
\mathrm{match}_i \propto \mathrm{Re}\left(\langle \sigma_+^i\, a\rangle e^{-i\Phi}\right),
\end{equation}
?????? ?????? ??????????? ??????????????? ?????? ???????? ?????????.

\paragraph{??????????? ??????????.}
????????? ????????? ????? ???? ??????????,
\begin{equation}
\mathcal{D}=1-\left|\langle a\rangle\right|.
\end{equation}
????? ?????? ????? ??????????:
\begin{equation}
\dot S_i \leftarrow \dot S_i - \gamma_{coh}(1-\mathcal{D})S_i.
\end{equation}
?? ???????????? ?????????????? ??? ???????? ????????????, ?????????? ????????????? QRC.

\paragraph{?????? ??? ????.}
??????? ? ?????????: (i) ?????? ???????? (???? ??????????), (ii) ????????????? ????? ?????? ??????? ?????, (iii) ??????????? ??????????. ?????? ??? ???? ????? ?? ???????????? ????????? ??? ??????? ???????.
\selectlanguage{english}

\section*{Open Additions}
Future work: explicit lag estimation, adaptive group-level delays, and hierarchical QRC (local + global).

\end{document}
