\documentclass[12pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[russian,english]{babel}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{geometry}
\usepackage{hyperref}
\geometry{margin=2.5cm}

\title{
    \textbf{Meta-Stable Architectures}\\
    \Large Volume I (v2): Foundations of Metastable Dynamics in Adaptive Systems
}
\author{Symbiosis: Katia \& GPT-5 (Jippi)}
\date{2026}

\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}

\begin{document}
\maketitle

\begin{abstract}
We formalize metastable dynamics for adaptive dissipative systems. In v2, the regulatory variable is modeled with bounded-input stable dynamics to remove boundedness inconsistencies. We prove forward completeness and global boundedness of the full state under explicit assumptions.
\end{abstract}

\selectlanguage{russian}
\begin{abstract}
В версии v2 формализована метастабильная динамика адаптивных диссипативных систем с устранением противоречий по ограниченности. Регуляторная переменная задается устойчивой динамикой с ограниченным входом. Доказываются прямая полнота и глобальная ограниченность полного состояния.
\end{abstract}
\selectlanguage{english}

\section*{Version Notes (v2)}
\begin{itemize}
    \item Corrected the regulatory-state dynamics to support boundedness claims.
    \item Revised Theorem 1 to match explicit assumptions.
    \item Completed Lyapunov dissipation estimates.
    \item Added quantitative metastability definition via exit-time.
\end{itemize}

\section{System}
Let $a(t)\in\mathbb{R}^n$ be the primary state and $\Phi(t)\in\mathbb{R}_+$ the regulatory activity:
\begin{align}
\dot{a} &= -\nabla V(a)-\lambda S(a,\Phi), \label{eq:a}\\
\dot{\Phi} &= -\kappa \Phi + u(a), \label{eq:phi}
\end{align}
where $\lambda,\kappa>0$, $V:\mathbb{R}^n\to\mathbb{R}$.

\section{Assumptions}
\begin{itemize}
    \item[(A1)] $V\in C^2(\mathbb{R}^n)$ and $V(a)\to\infty$ as $|a|\to\infty$.
    \item[(A2)] There exist $c_1>0$, $R>0$ such that $|\nabla V(a)|\ge c_1|a|$ for $|a|>R$.
    \item[(A3)] $S(a,\Phi)$ is locally Lipschitz in $(a,\Phi)$ and uniformly bounded: $|S(a,\Phi)|\le C_S$.
    \item[(A4)] $u(a)$ is bounded: $|u(a)|\le U_{\max}$.
\end{itemize}

\section{Global Boundedness and Forward Completeness}
\begin{theorem}
Under (A1)--(A4), for every initial condition $(a_0,\Phi_0)$ the solution of \eqref{eq:a}--\eqref{eq:phi} exists for all $t\ge 0$ and satisfies
\[
\sup_{t\ge0}\left(|a(t)|+|\Phi(t)|\right)<\infty.
\]
\end{theorem}

\begin{proof}[Proof sketch]
Take $V(a)$ as Lyapunov candidate for the $a$-subsystem:
\[
\dot V = \nabla V\cdot\dot a = -|\nabla V|^2-\lambda \nabla V\cdot S(a,\Phi)
\le -|\nabla V|^2+\lambda C_S|\nabla V|.
\]
Using Young's inequality:
\[
\dot V\le -\frac12|\nabla V|^2+\frac12\lambda^2C_S^2.
\]
By (A2), outside a ball, $|\nabla V|^2\ge c_1^2|a|^2$, so $\dot V\le -\alpha |a|^2+\beta$ for some $\alpha,\beta>0$. Hence $a(t)$ enters a bounded absorbing set and cannot blow up in finite time.

For \eqref{eq:phi},
\[
\Phi(t)=e^{-\kappa t}\Phi_0+\int_0^t e^{-\kappa(t-s)}u(a(s))\,ds,
\]
thus
\[
|\Phi(t)|\le |\Phi_0|+\frac{U_{\max}}{\kappa}.
\]
So $(a,\Phi)$ is globally bounded and the solution is forward complete.
\end{proof}

\section{Metastable Trajectory}
\begin{definition}
Let $A$ be a local attractor and $U(A)$ its bounded neighborhood. A trajectory is metastable if:
\begin{enumerate}
    \item there exists $T>0$ such that $a(t)\in U(A)$ for $t\in[0,T]$;
    \item for an admissible bounded perturbation, the exit time
    \[
    \tau_U=\inf\{t>0:\, a(t)\notin U(A)\}
    \]
    is finite.
\end{enumerate}
\end{definition}

\section*{DOI}
Volume I v1 DOI: \href{https://doi.org/10.5281/zenodo.18815179}{10.5281/zenodo.18815179}.
This v2 supersedes v1 for formal mathematical use.

\end{document}
